\subsection{Language} \label{sec:Language}

We want to make an implementation of the Haskell programming
language~\cite{marlow2010haskell} with unlockable sets of features.  Haskell is
a declarative programming language suitable and often preferred for use as the
first taught language in computer science courses~\cite{dijkstra2001members}.
It is regarded as an elegant programming language that allows writing programs
in many paradigms including functional and imperative.

With unlockable features we mean that a student may start out with a language
with very few features turned on. As a result, we can provide very relevant
error messages -- tailored to their level of expertise -- that do not mention
features that they have not yet encountered. As the teaching progresses, we can
turn on increasingly advanced features of the language.  Since we are using an
already established, general-purpose language it means that the knowledge
gained using our implementation transfers to ordinary implementations of
Haskell, and by having unlockable features we avoid the problems associated
with using a general-purpose language for teaching.

\subsection{Integrated development environment} \label{sec:IDE}

To solve the problem of installation difficulties, we propose to provide an
online IDE for our language.  This means that a student can simply point their
web browser to a website providing the IDE and immediately start writing and
running code in the browser; the barrier to entry is extremely low. It also
means that we can provide the same up-to-date version of the language and its
libraries to everyone.

An online IDE opens the door to many features that will help the students to
learn or just make programming exciting:
\begin{description}
  \item[Collaborating]
    Students can form groups and collaboratively write code.  Since the IDE is
    online they can do this even if they are not physically in the same
    location, and they can edit the same file simultaneously.
  \item[Sharing]
    Students can easily share their code and applications with friends and
    classmates.
  \item[Writing games] 
    We can provide libraries to write games that run in the browser, and
    provide easy access to art assets in the IDE. Games are fun and rewarding
    to hack, which should help to increase the adoption of the language. It
    also lets us focus development, and sets us apart from other educational
    programming languages.
  \item[Learning]
    Teachers can provide learning material with exercises that students can do
    and have automatically marked directly in the IDE. Language features (or
    even game assets) can be unlocked as the students progress through the
    exercises. It is possible to envision branching capabilities in the
    exercise machinery, so that the user could be prompted for what they would
    like to learn next.  Throw some learning analytics at this baby, and we
    have informed suggestions as to what to learn next as well.
\end{description}

\subsection{Academic use}

We have already touched on a few uses of our language and IDE in the academic
sector, but there are many more possibilities.

Our IDE could be specialised for programming courses, with virtual classrooms
where students and tutors could meet to discuss the course and its contents.
Teachers could author their own custom curricula and exercises and we could
integrate the IDE with their learning management system.

By providing different libraries and toggling different sets of features of the
language it could be customised to suit many diverse kinds of courses at
different levels in the education system, be it programming fundamentals at the
university level or game development for kids.

It is likely that teachers would be very interested in grading systems, and 
integration with learning management systems. Generated reports would be 
useful for students and teachers both.

We should try to get some schools or computing groups to use our environment 
as we develop it, to get useful feedback. Our background in academia might
help us finding interested candidates.

\subsection{Commercialisation}

We would like to keep the basic IDE gratis for everyone, but we still have
several ideas that make the idea commercially viable:

\begin{description}
  \item[Premium features]
    The classroom features could for instance be a premium feature that schools
    and universities would have to pay for.
  \item[Art assets]
    Art assets for use in games and applications could be sold directly in the
    IDE. Or, we could integrate with existing art asset services. Users should 
    be allowed to gratis upload their own art assets as well.
  \item[Application store]
    Advanced users could be allowed to publish and sell their creations in an 
    associated application store in return for a portion of their revenue. 
    Teachers could sell course material as well.
\end{description}

\subsection{Societal benefits}
Whilst user interface designers keep telling us that user interface design is 
constantly improving, there is no denying the inherent complexity of a modern 
day computer. In a society where only software developers understand the basic 
science of a computer program, unenlightened computer users are left helpless. 
This is problematic in the case of proprietary software, where computer users 
are slaves to the subjugation of the power elite created by these software 
developers -- and worst of all, they often don't even realise it. Learning to 
program a computer will as a side-effect make computer users more aware of how 
computers actually work, which in turn gives them the foundation for 
self-educating further. Moreover, understanding the basics of computer 
programming means understanding what source code is, and why it must be free 
for computer users themselves to be free. Illuminating users in this manner is 
among the zenith of societal contribution in computer science.

Another aspect to consider is that teaching computer programming in a 
principled way is The Right Thing. Computer programmers today are typically 
either lone self-taught hackers, or computer science students. Both of these 
groups suffer setbacks from learning to program through pedagogically unsound 
tools. Others never make it to an enlightened state at all. Society at large 
is starting to take this seriously. Computing at 
School\footnote{\url{http://www.computingatschool.org.uk/}} has been 
successful in improving national school curricula in the UK, and in Norway we 
have Lær Kidsa Koding\footnote{\url{http://www.kidsakoder.no/}}. We believe
that our language could have a profound impact in these circles.

\subsection{Timetable}
Having unlockable subsets of the language makes it natural to develop the 
language iteratively. We will design the first stage, the basic level of the 
language, very meticulously, and merely delineate future stages. Then, as 
development goes on, we will continue this trend of meticulousness for 
short-term goals, and nice academic hand-waving of the future. It's nice our 
degrees weren't for naught.

After around 15 months of work, we should have all the complicated language 
design decisions mostly out of the way (though subject to change). A 
preliminary implementation of the unlocking concept should be in place, and 
the basic first level of the language, along with some basic libraries, should 
all be available. We should also have a third-party IDE connected to our 
language, for proof-of-concept. Finally, we should also have delineated some 
of the immediate next levels.
